Deep Probabilistic Models for Camera Geo-Calibration

### 2018-08-24 chapters/introduction/introduction.tex
background change

\begin{chapquote}{Morgan Freeman in {\em Transcendence (2014 film)}
\\ \todo{don't laugh.}}
  ``Geo-locating suspects in real time? I've never seen anything like
it.''
\end{chapquote}

@@

For many applications, knowing when, where, and in which direction a
picture was taken is important to understand the world. However, most
images does not carry such information, thus we need to develop
algorithms to dig them up.  We refer to the task of estimating these
properties as {\em geo-calibration}.
Most recent work of geo-calibration focus on deterministic systems, 
which have some drawbacks: 1) they can not model the inherent
uncertainties from images of ambiguous scenes, and 2) they can not be
used by applications that take probabilistic inputs. To address
these problems, we propose to build probabilistic models for camera
geo-calibration.

We also show that learning to geo-calibrate a camera is helpful for
high-level feature extracting. Feature engineering is essential for
image understanding. In recent years, learning based approaches for
high-level feature extracting like deep neural nets have drawn a lot
of attention due to the fact that they do not need any expert
knowledge about the target data. However, most of these feature
learning methods heavily relies on manually labeled data. In our
thesis, we presented alternative ways to learn high-level image
representations when the manually labeled data is either insufficient
or absent.

\section{Background}

\subsection{Geo-Calibration of Outdoor Images}
Automatic image geo-calibration of outdoor images continues to grow in
importance as a direct result of the increasing amount of imagery
available via the Internet.  Solving this problem is of great value
for a wide variety of fields, with potential applications ranging from
the forensic sciences~\cite{stylianou13jane} to environmental
monitoring~\cite{zhang2012mining}.  Conceptually the geo-calibration
task includes: identifying the camera pose ({\em pose estimation}),
the camera geographic location ({\em geo-localization}), and the time
when the query image was captured ({\em time estimation}).

@@

\item \textbf{Pose Estimation:}
Li et al.~\cite{li2012worldwide} exploit geo-registered 3D points
clouds to estimate camera pose.
Vo \etal~\cite{vo2016localizing} propose a geo-localization network
that can regress the geo-orientation angle of the camera.
Agarwal \etal~\cite{agarwal2015metric} keep track of the camera pose
changes by matching SIFT~\cite{lowe1999object} feature points detected
between the input image and Google Street View panoramas.
%
Horizon line detection is an essential step for some pose
estimation algorithms, because horizon lines are closely related to
camera pitch and roll angles.
%If the camera focal length is known, one can compute the camera pitch
%and roll angle from the horizon line.
Collins and Weiss~\cite{unitsphere1990} formulate horizon line
detection as a statistical estimation problem on the Gaussian Sphere,
which is similar to the geometry we use.  More recent work has
explored the use of dual space~\cite{alignment2014,dualspace2013}
representations. Among the clustering-based approaches, Xu et
al.~\cite{kitware2013} improve this pipeline by introducing a new
point-line consistency function that models errors in the line segment
extraction step.
\newline

\item \textbf{Geo-Localization:}
Extracting location-dependent features from
image data has drawn a great detail of attention from the vision
community~\cite{jacobs07geolocate, jacobs11geolocate,
jacobs08geoorient}. The common trend amongst these methods is that
they take advantage of a large dataset of geo-referenced images. Hays
and Efros~\cite{hays2008im2gps} use a data-driven scene matching
approach to localize a query image using a large dataset of geo-tagged
images.  Doersch \etal~\cite{doersch2012what} extract
location-dependent features that capture the relative appearance
differences of large cities.  Lin et al.~\cite{lin2013cross} localize
a ground-level image by learning the relationship between pairs of
ground and aerial images \footnote{In the context of our thesis, we
do not distinguish between the satellite imagery and the aerial
imagery} of the same location. Other techniques focus
on urban environments and infer location using local image
descriptors~\cite{schindler2008detecting,snavely2006photo}.
Many other cues exist, such as the
skyline~\cite{baatz2012large,ramalingam2009geolocalization}, sky
appearance~\cite{lalonde2010sun,workman2014rainbow}, and
shadows~\cite{junejo2008estimating,wu2010geo}.
\newline

\item \textbf{Time Estimation:}
Matzen and Snavely~\cite{matzen2014scene} predict
timestamps for photos by matching against a time-varying
reconstruction of a scene.  Hill~\cite{hill1994elephant} estimate the
time of the day by measuring the light intensity profiles captured by
cameras. Methods are proposed to date the
year book photos by analyzing human
appearance~\cite{salem2016face2year,ginosar2015century}. Lee
\etal~\cite{linking2015iccp} find visual paterns in the buildings,
relate them to certain time periods.
\newline

>>>>

\begin{chapquote}{Morgan Freeman in {\em Transcendence (2014 film)}}
  ``Geo-locating suspects in real time? I've never seen anything like
it.''
\end{chapquote}

@@

For many applications, knowing when, where, and in which direction a
picture was taken is important to understand the world. However, most
images does not carry such information. Our thesis focuses on
developing algorithms to estimate this information.
We refer to the task of estimating these properties as {\em
geo-calibration}.
Most recent work of geo-calibration focus on deterministic systems
\todo{citations}, 
which have some drawbacks: 1) they can not model the inherent
uncertainties from images of ambiguous scenes, and 2) without a proper
modeling of the uncertainty, subsequent processing can yield overly
confident predictions. To address these problems, we propose to build
probabilistic models for camera geo-calibration.

We also show that learning to geo-calibrate a camera allows us to
implicitly learn to understand the content of the scene.
As an essential part of scene understanding, how to extract image
features is an important problem being studied for decades.  In
recent years, learning based approaches for image feature extraction
like convolutional neural networks have drawn a lot of attention due
to the fact that they do not need expert knowledge about the target
data. However, most of these feature learning methods require large
quantity of manually labeled data sets.  In this thesis, we presented
alternative ways to learn high-level image features when the
manually labeled data is either insufficient or absent.

\section{Background}
For a better understanding of our work, before the discussion of the
main topics, we would like to introduce some related computer vision
concepts in this section.

\subsection{Geo-Calibration of Outdoor Images}
Automatic geo-calibration of outdoor images continues to grow in
importance as a direct result of the increasing amount of imagery
available via the Internet. Solving this problem is of great value
for a wide variety of fields, with potential applications ranging from
the forensic sciences~\cite{stylianou13jane} to environmental
monitoring~\cite{zhang2012mining}. Conceptually the geo-calibration
task includes: identifying the camera pose ({\em pose estimation}),
the camera geographic location ({\em geo-localization}), and the time
when the query image was captured ({\em time estimation}).

@@

\item \textbf{Pose Estimation:}
Pose estimation includes estimating the camera yaw, pitch, and roll
angles, $(\theta_{yaw}, \theta_{pitch}, \theta_{roll})$ from a
image. In the context of geo-calibration, the yaw angle
$\theta_{yaw}$ is referred to as the geographic orientation angle of
the camera.
%
Li et al.~\cite{li2012worldwide} exploit geo-registered 3D points
clouds to estimate camera pose.
Vo \etal~\cite{vo2016localizing} propose a geo-localization network
that can regress the geo-orientation angle of the camera.
Agarwal \etal~\cite{agarwal2015metric} keep track of the camera pose
changes by matching SIFT~\cite{lowe1999object} feature points detected
between the input image and Google Street View panoramas.
%
Horizon line detection is an essential step for some pose
estimation algorithms, because horizon lines are closely related to
camera pitch and roll angles.
%If the camera focal length is known, one can compute the camera pitch
%and roll angle from the horizon line.
Collins and Weiss~\cite{unitsphere1990} formulate horizon line
detection as a statistical estimation problem on the Gaussian Sphere,
which is similar to the geometry we use.  More recent work has
explored the use of dual space~\cite{alignment2014,dualspace2013}
representations. Among the clustering-based approaches, Xu et
al.~\cite{kitware2013} improve this pipeline by introducing a new
point-line consistency function that models errors in the line segment
extraction step.
\newline

\item \textbf{Geo-Localization:}
The task of geo-localization estimates the geographic location, $(lat,
lon)$, of the camera from the input image or video frames.
%
As an important step of geo-localization, extracting
location-dependent features from image data has drawn a great detail
of attention from the vision community~\cite{jacobs07geolocate,
jacobs11geolocate, jacobs08geoorient}. The common trend amongst these
methods is that they take advantage of a large dataset of
geo-referenced images. Hays and Efros~\cite{hays2008im2gps} use a
data-driven scene matching approach to localize a query image using a
large dataset of geo-tagged images.  Doersch
\etal~\cite{doersch2012what} extract location-dependent features that
capture the relative appearance differences of large cities.  Lin et
al.~\cite{lin2013cross} localize a ground-level image by learning the
relationship between pairs of ground and aerial images \footnote{In
the context of our thesis, we do not distinguish between satellite
imagery and aerial imagery} of the same location. Other techniques
focus on urban environments and infer location using local image
descriptors~\cite{schindler2008detecting,snavely2006photo}.  Many
other cues exist, such as the
skyline~\cite{baatz2012large,ramalingam2009geolocalization}, sky
appearance~\cite{lalonde2010sun,workman2014rainbow}, and
shadows~\cite{junejo2008estimating,wu2010geo}.
\newline

\item \textbf{Time Estimation:}
The goal of time estimation is to estimate the time when the input
image was captured. Depending on the application, the scale of a
time estimator outputs may range from year to hour.
%
Matzen and Snavely~\cite{matzen2014scene} predict
timestamps for photos by matching against a time-varying
reconstruction of a scene.  Hill~\cite{hill1994elephant} estimate the
time of the day by measuring the light intensity profiles captured by
cameras. Methods are proposed to date the
year book photos by analyzing human
appearance~\cite{salem2016face2year,ginosar2015century}. Lee
\etal~\cite{linking2015iccp} find visual paterns in the buildings,
relate them to certain time periods.
\newline

### 2018-08-24 dissertation.tex
background change

\abstract
{ \SingleSpacing
The ultimate goal of image understanding is to transfer visual
images into numerical or symbolic descriptions of the scene
that are helpful for decision making.
%
%In general, understanding images is not trivial for machines.
%Under different scenarios, researchers divide image
%understanding into different computer vision tasks. For example,
%pedestrian detection is essential to autonomous driving, and image
%retrieval to an image searching engine.
%
Knowing when, where, and in which direction picture was taken, the
task of geo-calibration makes it possible to use imagery to understand
the world and how it changes in time. Current models for
geo-calibration are mostly deterministic, which fails to model the
the inherent uncertainties when the scene of the image is ambiguous.
Furthermore, the outputs of deterministic system can not be used by
algorithms that take probabilistic inputs. To address these
limitations, we propose a probabilistic model for camera
geo-calibration using deep neural networks.
%
While our primary contribution is geo-calibration, We also show that
learning to geo-calibrate a camera is helpful for high-level feature
extraction.
%
%Our discoveries can provide researchers new insights for both camera
%geo-calibration and DNN training.
}
\keywords{computer vision, geo-calibration, deep neural networks}

>>>>

\abstract
{ \SingleSpacing
The ultimate goal of image understanding is to transfer visual
images into numerical or symbolic descriptions of the scene
that are helpful for decision making.
Knowing when, where, and in which direction a picture was taken, the
task of geo-calibration, makes it possible to use imagery to understand
the world and how it changes in time. Current models for
geo-calibration are mostly deterministic, which in many cases fails to
model the inherent uncertainties when the image content is
ambiguous. Furthermore, without a proper modeling of the uncertainty, subsequent
processing can yield overly confident predictions.
To address these limitations, we propose a probabilistic model for
camera geo-calibration using deep neural networks.
While our primary contribution is geo-calibration, we also show that
learning to geo-calibrate a camera allows us to implicitly learn to
understand the content of the scene.
}
\keywords{computer vision, geo-calibration, deep neural networks}
